---
title: "On the Topic of the Exponential Distribution with respect to the Central Limit Theorem in R"
output:
  pdf_document: default
  html_document: default
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "Evan Falcone"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This data analysis aims to investigate the Exponential distribution with respect to the Central Limit Theorem under certain constraints. We will use lambda = 0.2 for all simulations. We compare the distribution of averages of 40 exponentials over 1000 simulations.

## Simulations

Start by initializing the common variables (lambda, exponentials) and setting a seed:

```{r initial, echo=TRUE}
set.seed(1337)
lambda <- 0.2
exponentials <- 40
```

Run simulations with the initialized variables:

```{r simulation, echo=TRUE}
simMean <- NULL
for(i in 1:1000) {simMean <- c(simMean, mean(rexp(exponentials, lambda)))}
```

## Comparison: Sample Mean to Theoretical Mean

### Sample Mean

We compute the sample mean by taking the mean of the simulations:

```{r sampMean, echo=TRUE}
mean(simMean)
```

### Theoretical Mean

The theoretical mean of an exponential distribution is given by lammbda ^ -1:

```{r theoMean, echo=TRUE}
lambda^-1
```

### Comparison

The difference between the simulations' sample mean and the theoretical mean of the exponential distribution is small (order of magnitude = 10^-2):

```{r compMean, echo=TRUE}
abs(mean(simMean)-(lambda^-1))
```

## Comparison: Sample Variance to Theoretical Variance

### Sample Variance

We compute the sample variance by finding the variance of the simulations:

```{r sampVar, echo=TRUE}
var(simMean)
```

### Theoretical Variance

The theoretical variance of an exponential distribution is given by (lammbda * sqrt(n))^-2:

```{r theoVar, echo=TRUE}
(lambda * sqrt(exponentials))^-2
```

### Comparison

The difference between the simulations' sample variance and the theoretical variance of the exponential distribution is small (order of magnitude = 10^-2):

```{r compVar, echo=TRUE}
abs(var(simMean)-(lambda * sqrt(exponentials))^-2)
```

## Distribution

The following is a density histogram plot of the 1000 simulations made using our initialized parameters lambda = 0.2 and seed 1337. We overlay a normal distribution that has a mean == mean(simMean) (i.e. lamda^-1) and variance == var(simMean) (i.e. (lambda * sqrt(exponentials))^-2), the theoretical normal distribution for the simulations we made. Graphically, the two distributions appear to bear a resemblance:

```{r densityPlot, echo=TRUE}

hist(simMean, prob=TRUE, col="lightblue", main="Mean distribution for rexp()", breaks=20)
lines(density(simMean), lwd=3, col="blue")
```

